The Spatial and Temporal Domains of Modern Ecology

### 2017-05-14 Rmd/space-time.Rmd
Paper revised through end of text describing Figure 2.

aaxis2 <- cbind.data.frame(
  c(0.01 / 10000, 0.1 / 10000, 1 / 10000, 10 / 10000, 100 / 10000, 1000 / 10000,
    1, 10, 10^2, 10^3, 10^4, 10^5, 10^6, 10^7, 10^8, 10^9, 10^10))
aaxis2 <- data.table(aaxis2)
setnames(aaxis2, names(aaxis2), "area")
aaxis2[, logarea := log10(area)]
```

@@

# stats
# bootp_dt
# axisl <- list(aaxis1$logres, c(-6, aaxis2$logarea), taxis1$logdays,
#               taxis2$logdays)
# hdatstat <- lapply(1:4, function(x) {
#   h <- hist(hdat[[x]], breaks = brksl[[x]], plot = FALSE)
#   h$density <- h$counts / sum(h$counts) * 100
#   h
# })
```

Statistics associated with Figure 1: percentage of observations falling within different scale ranges, and associated with different observational methods
```{r, eval = FALSE}
# res
# hdatst <- hdatstat[[1]]$density
hdatst <- hdatstat_bt$res$mu
l <- length(hdatst)
sum(hdatst[4:6])  # 53 % between 10cm^2 and 1 m^2
sum(hdatst[1:6])  #  67% < 1m^2
sum(hdatst[7:10]) # 24 % between 1 m^2 and 1 ha
sum(hdatst[11:(l)]) # 9 % 1-10000 ha
# dat[plot_res >= 10000^2]

@@

# extent
# hdatst <- hdatstat[[2]]$density
hdatst <- hdatstat_bt$ext$mu
l <- length(hdatst)
sum(hdatst[1:2])  # 36 % < 1 m^2
sum(hdatst[1:6])  # 82% < 1 ha
sum(hdatst[1:7])  # 86% < 10 ha
sum(hdatst[1:8])  # 90 % < 100 ha
sum(hdatst[1:9])  # 92 % < 1000 ha
sum(hdatst[8:9])  # 6 % 10-1000 ha
sum(hdatst[8:l])  # 14% > 10 ha
sum(hdatst[10:l]) # 8.2 % covered an area > 1000 ha
sum(hdatst[12:l]) # 5.1 % covered an area > 100,000 ha
sum(hdatst[13:l]) # 3.9 % covered an area > 1,000,000 ha

# length(which(hdat$ext < -4)) / length(hdat$ext)
# length(which(hdat$ext > 3)) / length(hdat$ext)
# length(which(hdat$ext > 1 & hdat$ext <= 3)) / length(hdat$ext)
# which(hdatstat[[2]]$breaks == log10(1000000))
# length(hdatstat[[2]]$counts[7:8])
# hdatstat[[2]]$mids[12]

# interval
# hdatst <- hdatstat[[3]]$density
hdatst <- hdatstat_bt$int$mu
l <- length(hdatst)
sum(hdatst[l])  # 36% are once-offs
sum(hdatst[1:3])  # 15 % < daily
sum(hdatst[1:4])  # 25 % < weekly
sum(hdatst[4:5])  # 21% daily up to monthly
sum(hdatst[6])  # 19% monthly up to yearly
sum(hdatst[7:8]) # 8.2 % yearly to decadal

# duration
# hdatst <- hdatstat[[4]]$density
hdatst <- hdatstat_bt$dur$mu
l <- length(hdatst)
sum(hdatst[1:3])  # 71 % up to 1 day
sum(hdatst[4:5])  # 17% covered between 1 day and 1 month
sum(hdatst[6])  # 7.3% covered between 1 month and 1 year
# sum(hdatstat[[4]]$density[4:6])  # 26% covered between 1 day and 1 year
sum(hdatst[7:10]) # 4.5% year to century

# datf[, plot(log10(n_sites), log10(plot_res))]
# dat[like(study_type, "field"), plot(log10(n_sites), log10(plot_res))]
datf[like(study_type, "field"), plot(log10(plot_res), log10(act_ext), 
                                     ylim = c(-10, 10))]
datf[like(study_type, "remote"), 
     points(log10(plot_res), log10(act_ext), pch = 20, col = "red")]
datf[like(study_type, "automated"), 
     points(log10(plot_res), log10(act_ext), pch = 20, col = "blue")]
lines(aaxis1$logres, aaxis2$logarea[-c(16:17)] )
datf[, .N / nrow(datf) * 100, by = st]

datf[, plot(log10(t_btwn_samp), log10(act_dur), pch = 20)]
datf[which(hdat[[3]] < -1.380211), 
     points(log10(t_btwn_samp), log10(act_dur), pch = 20, col = "blue")]
datf[like(study_type, "automated"), 
     points(log10(t_btwn_samp), log10(act_dur), pch = 20, col = "red")]

# types of observations
# unique(datf$study_type)
datf[like(study_type, "field"), .N] / nrow(datf)  # 80% field studies
datf[like(study_type, "auto"), .N] / nrow(datf) # 12 % automated
datf[like(study_type, "auto") & log10(t_btwn_samp) < -1.380211, .N] / 
  length(which(hdat[["int"]] < -1.380211))  # 78 % high frequency obs automated 
datf[like(study_type, "remote|other"), .N] / nrow(datf) # 7 % remote
datf[like(study_type, "paleo"), .N] / nrow(datf) * 100 # <1 % paleo

@@

# pdf("paper/figures/res_v_extent.pdf", width = 7, height = 2.5)
pdf("paper/figures/kde45.pdf", width = 7, height = 6)
par(mfrow = c(2, 2), mar = c(6, 6, 0.3, 3), oma = c(0, 0, 0.5, 1), 
    mgp = c(2, 0.25, 0))
# resolutions
# image(rres3, col = rev(terrain.colors(30)), breaks = brks,
image(rres3, axes = FALSE, xlab = "", ylab = "", #rev(terrain.colors(cuts))) 
      col = cols)
axis(1, at = taxis1$logdays[-c(5)], labels = tlab1[-c(5)], 
     las = 2, tcl = -0.2, cex.axis = cxa)
axis(2, at = aaxis1$logres[-1], labels = alab1[-1], las = 2, tcl = -0.2, 
     cex.axis = cxa)
mtext("Sampling interval", 1, cex = cxa, line = ll)
mtext("Spatial resolution", 2, cex = cxa, line = ll)
brks <- brkfun(rres3, 0.01, cuts)  # breaks
aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
              cex.axis = cxa, tcl = -0.1)
plot(rres3, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     #legend.shrink = 0.9, breaks = brks, col = rev(terrain.colors(30)))
     legend.shrink = 0.9, col = cols)#inferno(cuts))
     #col = rev(terrain.colors(cuts)))
mtext("A", side = 3, line = -1, cex = 0.8, adj = 0.05, col = "white")
# for(i in 1:length(stps)) {
#   points(resdat3$x[stps[[i]]], resdat3$y[stps[[i]]], pch=pchs[[i]], cex=cexs[i])
# }

# extents
# image(rext3, col = rev(terrain.colors(30)), axes = FALSE, xlab = "", 
image(rext3, axes = FALSE, xlab = "", ylab = "", col = cols)
axis(1, at = taxis2$logdays[-5], labels = tlab2[-5], las = 2,
     cex.axis = cxa, tcl = -0.2)
axis(2, at = aaxis2$logarea, labels = alab2, las = 2, cex.axis = cxa,
     tcl = -0.2)
brks <- brkfun(rext3, 0.01, cuts)  # breaks
aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
              cex.axis = cxa, tcl = -0.1)
mtext("Temporal duration", 1, cex = cxa, line = ll)
mtext("Spatial extent", 2, cex = cxa, line = ll)
plot(rext3, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     legend.shrink = 0.9, col = cols)
mtext("B", side = 3, line = -1, cex = 0.8, adj = 0.05, col = "white")

# spatial
image(sres, col = cols, axes = FALSE, xlab = "", ylab = "") 
axis(1, at = aaxis1$logres, labels = alab1, las = 2, tcl = -0.2, cex.axis = cxa)
axis(2, at = aaxis2$logarea, labels = alab2, las = 2, tcl = -0.2, 
     cex.axis = cxa)
mtext("Spatial resolution", 1, cex = cxa, line = ll)
mtext("Spatial extent", 2, cex = cxa, line = ll)
brks <- brkfun(sres, 0.01, cuts)  # breaks
aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
              cex.axis = cxa, tcl = -0.1)
plot(sres, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     legend.shrink = 0.9, col = cols)
mtext("C", side = 3, line = -1, cex = 0.8, adj = 0.05, col = "white")
# for(i in 1:length(stps)) {
#   points(spatdat$x[stps[[i]]], spatdat$y[stps[[i]]], pch=pchs[[i]], cex=cexs[i])
# }

# temporal
image(tres, axes = FALSE, xlab = "", ylab = "", col = cols)
axis(1, at = taxis1$logdays[-c(5)], labels = tlab1[-c(5)], 
     las = 2, tcl = -0.2, cex.axis = cxa)
axis(2, at = taxis2$logdays[-5], labels = tlab2[-5], las = 2,
     cex.axis = cxa, tcl = -0.2)
mtext("Sampling interval", 1, cex = cxa, line = ll)
mtext("Temporal duration", 2, cex = cxa, line = ll)
# brks <- brkfun(tres, 0.02, 100)  # breaks
brks <- brkfun(tres, 0.01, cuts)  # breaks
aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
              cex.axis = cxa, tcl = -0.1)
plot(tres, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     legend.shrink = 0.9, col = cols)
mtext("D", side = 3, line = -1, cex = 0.8, adj = 0.05, col = "white")
dev.off()

# image(rres3, col = rev(terrain.colors(100)))
# points(resdat3$x, resdat3$y, pch = 20, cex = 0.5)
# image(sres, col = rev(terrain.colors(100)))
# points(spatdat$x, spatdat$y, pch = 20, cex = 0.5)
```

>>>>

aaxis2 <- cbind.data.frame(
  c(0.01 / 10000, 0.1 / 10000, 1 / 10000, 10 / 10000, 100 / 10000, 1000 / 10000,
    1, 10, 10^2, 10^3, 10^4, 10^5, 10^6, 10^7, 10^8, 10^9, 10^10))
aaxis2 <- data.table(aaxis2)
setnames(aaxis2, names(aaxis2), "area")
aaxis2[, logarea := log10(area)]
```
```{r, eval = FALSE, echo = FALSE}
save(aaxis1, aaxis2, taxis1, taxis2, alab1, alab2, tlab1, tlab2, 
     file = fp(p_dat, "derived/dimbreaks.rda"))
```

@@

# stats
# bootp_dt
# axisl <- list(aaxis1$logres, c(-6, aaxis2$logarea), taxis1$logdays,
#               taxis2$logdays)
# hdatstat <- lapply(1:4, function(x) {
#   h <- hist(hdat[[x]], breaks = brksl[[x]], plot = FALSE)
#   h$density <- h$counts / sum(h$counts) * 100
#   h
# })
```
```{r, echo = FALSE, eval = FALSE}
save(hdat, hdatstat_bt, file = fp(p_dat, "derived/hdatstat.rda"))
```

Statistics associated with Figure 1: percentage of observations falling within different scale ranges, and associated with different observational methods
```{r, echo = FALSE}
load(fp(p_dat, "derived/hdatstat.rda"))
load(fp(p_dat, "derived/dimbreaks.rda"))
```
```{r}
# res
# sum(hdatst[4:6])  # 53 % between 10cm^2 and 1 m^2
# sum(hdatst[1:6])  #  67% < 1m^2
# sum(hdatst[7:10]) # 24 % between 1 m^2 and 1 ha
# sum(hdatst[11:(l)]) # 9 % 1-10000 ha
# dat[plot_res >= 10000^2]
hdatst <- hdatstat_bt$res$mu
l <- length(hdatst)
hds <- c("10cm^2 - 1 m^2", "<1m^2", "1 m^2-1 ha", "1-10000 ha")
inds <- list(4:6, 1:6, 7:10, 11:l)
sapply(1:length(inds), function(x) {
  v <- round(sum(hdatst[inds[[x]]]), 1)
  names(v) <- hds[x]
  v
})

@@

# extent
# hdatst <- hdatstat[[2]]$density
hdatst <- hdatstat_bt$ext$mu
l <- length(hdatst)
# sum(hdatst[1:2])  # 36 % < 1 m^2
# sum(hdatst[1:6])  # 82% < 1 ha
# sum(hdatst[1:7])  # 86% < 10 ha
# sum(hdatst[1:8])  # 90 % < 100 ha
# sum(hdatst[1:9])  # 92 % < 1000 ha
# sum(hdatst[8:9])  # 6 % 10-1000 ha
# sum(hdatst[8:l])  # 14% > 10 ha
# sum(hdatst[10:l]) # 8.2 % covered an area > 1000 ha
# sum(hdatst[12:l]) # 5.1 % covered an area > 100,000 ha
# sum(hdatst[13:l]) # 3.9 % covered an area > 1,000,000 ha
hds <- c("<1 ha", "<10 ha", "<100 ha", "<1000ha", "10-1K ha", ">10ha", "1K-1MK",
         ">1Kha", ">100Kha", ">1Mha")
inds <- list(1:6, 1:7, 1:8, 1:9, 8:9, 8:l, 10:12, 10:l, 12:l, 13:l)
sapply(1:length(inds), function(x) {
  v <- round(sum(hdatst[inds[[x]]]), 1)
  names(v) <- hds[x]
  v
})

# interval
# hdatst <- hdatstat[[3]]$density
hdatst <- hdatstat_bt$int$mu
l <- length(hdatst)
# sum(hdatst[l])  # 36% are once-offs
# sum(hdatst[1:3])  # 15 % < daily
# sum(hdatst[1:4])  # 25 % < weekly
# sum(hdatst[4:5])  # 21% daily up to monthly
# sum(hdatst[6])  # 19% monthly up to yearly
# sum(hdatst[7:8]) # 8.2 % yearly to decadal
hds <- c("norep", "<daily", "<weekly",  "day-mo", "mo-yr", "yr-dec")
inds <- list(l, 1:3, 1:4, 4:5, 6, 7:8)
sapply(1:length(inds), function(x) {
  v <- round(sum(hdatst[inds[[x]]]), 1)
  names(v) <- hds[x]
  v
})

# duration
# hdatst <- hdatstat[[4]]$density
hdatst <- hdatstat_bt$dur$mu
l <- length(hdatst)
# sum(hdatst[1:3])  # 71 % up to 1 day
# sum(hdatst[4:5])  # 17% covered between 1 day and 1 month
# sum(hdatst[6])  # 7.3% covered between 1 month and 1 year
# # sum(hdatstat[[4]]$density[4:6])  # 26% covered between 1 day and 1 year
# sum(hdatst[7:10]) # 4.5% year to century
hds <- c("<=daily", "day-mo", "mo-yr", "yr-dec", "yr-cent")
inds <- list(1:3, 4:5, 6, 7, 8:10)
sapply(1:length(inds), function(x) {
  v <- round(sum(hdatst[inds[[x]]]), 1)
  names(v) <- hds[x]
  v
})

# datf[, plot(log10(n_sites), log10(plot_res))]
# dat[like(study_type, "field"), plot(log10(n_sites), log10(plot_res))]
# datf[like(study_type, "field"), plot(log10(plot_res), log10(act_ext), 
#                                      ylim = c(-10, 10))]
# datf[like(study_type, "remote"), 
#      points(log10(plot_res), log10(act_ext), pch = 20, col = "red")]
# datf[like(study_type, "automated"), 
#      points(log10(plot_res), log10(act_ext), pch = 20, col = "blue")]
# lines(aaxis1$logres, aaxis2$logarea[-c(16:17)] )
# datf[, .N / nrow(datf) * 100, by = st]

# datf[, plot(log10(t_btwn_samp), log10(act_dur), pch = 20)]
# datf[which(hdat[[3]] < -1.380211), 
#      points(log10(t_btwn_samp), log10(act_dur), pch = 20, col = "blue")]
# datf[like(study_type, "automated"), 
#      points(log10(t_btwn_samp), log10(act_dur), pch = 20, col = "red")]

# types of observations
# unique(datf$study_type)
datf[, round(.N / nrow(datf) * 100, 2), by = study_type]
round(datf[study_type == "automated" & log10(t_btwn_samp) < -1.380211, .N] / 
        hdat[int < -1.380211, .N] * 100, 2)
```

@@

# pdf("paper/figures/res_v_extent.pdf", width = 7, height = 2.5)
pdf("paper/figures/kde45.pdf", width = 7, height = 6)
par(mfrow = c(2, 2), mar = c(6, 6, 0.3, 3), oma = c(0, 0, 0.5, 1), 
    mgp = c(2, 0.25, 0))
# resolutions
# image(rres3, col = rev(terrain.colors(30)), breaks = brks,
image(rres3, axes = FALSE, xlab = "", ylab = "", #rev(terrain.colors(cuts))) 
      col = cols)
axis(1, at = taxis1$logdays[-c(5)], labels = tlab1[-c(5)], 
     las = 2, tcl = -0.2, cex.axis = cxa)
axis(2, at = aaxis1$logres[-1], labels = alab1[-1], las = 2, tcl = -0.2, 
     cex.axis = cxa)
mtext("Interval", 1, cex = cxa, line = ll)
mtext("Resolution", 2, cex = cxa, line = ll)
brks <- brkfun(rres3, 0.01, cuts)  # breaks
aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
              cex.axis = cxa, tcl = -0.1)
plot(rres3, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     #legend.shrink = 0.9, breaks = brks, col = rev(terrain.colors(30)))
     legend.shrink = 0.9, col = cols)#inferno(cuts))
     #col = rev(terrain.colors(cuts)))
mtext("A", side = 3, line = -1, cex = 0.8, adj = 0.05, col = "white")
# for(i in 1:length(stps)) {
#   points(resdat3$x[stps[[i]]], resdat3$y[stps[[i]]], pch=pchs[[i]], cex=cexs[i])
# }

# extents
# image(rext3, col = rev(terrain.colors(30)), axes = FALSE, xlab = "", 
image(rext3, axes = FALSE, xlab = "", ylab = "", col = cols)
axis(1, at = taxis2$logdays[-5], labels = tlab2[-5], las = 2,
     cex.axis = cxa, tcl = -0.2)
axis(2, at = aaxis2$logarea, labels = alab2, las = 2, cex.axis = cxa,
     tcl = -0.2)
brks <- brkfun(rext3, 0.01, cuts)  # breaks
aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
              cex.axis = cxa, tcl = -0.1)
mtext("Duration", 1, cex = cxa, line = ll)
mtext("Extent", 2, cex = cxa, line = ll)
plot(rext3, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     legend.shrink = 0.9, col = cols)
mtext("B", side = 3, line = -1, cex = 0.8, adj = 0.05, col = "white")

# spatial
image(sres, col = cols, axes = FALSE, xlab = "", ylab = "") 
axis(1, at = aaxis1$logres, labels = alab1, las = 2, tcl = -0.2, cex.axis = cxa)
axis(2, at = aaxis2$logarea, labels = alab2, las = 2, tcl = -0.2, 
     cex.axis = cxa)
mtext("Resolution", 1, cex = cxa, line = ll)
mtext("Extent", 2, cex = cxa, line = ll)
brks <- brkfun(sres, 0.01, cuts)  # breaks
aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
              cex.axis = cxa, tcl = -0.1)
plot(sres, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     legend.shrink = 0.9, col = cols)
mtext("C", side = 3, line = -1, cex = 0.8, adj = 0.05, col = "white")
# for(i in 1:length(stps)) {
#   points(spatdat$x[stps[[i]]], spatdat$y[stps[[i]]], pch=pchs[[i]], cex=cexs[i])
# }

# temporal
image(tres, axes = FALSE, xlab = "", ylab = "", col = cols)
axis(1, at = taxis1$logdays[-c(5)], labels = tlab1[-c(5)], 
     las = 2, tcl = -0.2, cex.axis = cxa)
axis(2, at = taxis2$logdays[-5], labels = tlab2[-5], las = 2,
     cex.axis = cxa, tcl = -0.2)
mtext("Interval", 1, cex = cxa, line = ll)
mtext("Duration", 2, cex = cxa, line = ll)
# brks <- brkfun(tres, 0.02, 100)  # breaks
brks <- brkfun(tres, 0.01, cuts)  # breaks
aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
              cex.axis = cxa, tcl = -0.1)
plot(tres, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     legend.shrink = 0.9, col = cols)
mtext("D", side = 3, line = -1, cex = 0.8, adj = 0.05, col = "white")
dev.off()

# image(rres3, col = rev(terrain.colors(100)))
# image(rext3, col = rev(terrain.colors(100)))
# datf[, points(log10(eff_dur), log10(eff_ext), pch = 20, cex = 0.5)]
# datf[like(study_type, "remote"), {
#   points(log10(eff_dur), log10(eff_ext), pch = 20, cex = 0.5, col = "blue")
# }]
# points(resdat3$x, resdat3$y, pch = 20, cex = 0.5)
# image(sres, col = rev(terrain.colors(100)))
# datf[, points(log10(plot_res), log10(eff_ext), pch = 20, cex = 0.5)]
# datf[like(study_type, "remote"), {
#   points(log10(plot_res), log10(eff_ext), pch = 20, cex = 0.5, col = "blue")
# }]
# datf[like(study_type, "field"), {
#   points(log10(plot_res), log10(eff_ext), pch = 20, cex = 0.5, col = "red")
#   abline(lm(log10(plot_res) ~ log10(eff_ext)))
# }]
# points(spatdat$x, spatdat$y, pch = 20, cex = 0.5)
```

### 2017-05-14 paper/main.tex
Paper revised through end of text describing Figure 2.

We calculated these dimensions from 379 discrete observations (defined here as data collected from non-experimentally manipulated, or ``natural'' \cite{tilman_ecological_1989}, systems) reported within a 134 paper subset of 348 randomly selected articles (from 42,918 total). An additional 62 papers that were cited as the source of observations in the selected papers were also reviewed.

@@

In terms of resolution, the majority (67\%) of observations were collected in plots of $<$1 m$^2$ resolution, 24\% were collected within plots of 1 m$^2$ up to 1 ha, and the remaining 9\% in plots of $\geq$1 ha (Fig. 1A). The total spatial extent (the number of sampled sites multiplied by the spatial resolution) of 86\% of observations was $<$10 ha, while 36\% covered less than 1 m$^2$ (Fig. 1B). Only 14\% covered an extent $\geq$10 ha, with just 4\% spanning areas $\geq$1 million ha.

@@

\begin{figure}[!ht]
%\begin{wrapfigure}{c}{1\textwidth}
\includegraphics[width=1\textwidth]{figures/hists_bs.pdf}
\vspace{-0.15 cm}
\caption{Histograms of the spatial resolution (A) and extent (B), sampling interval (C) and temporal duration (D) of ecological observations collected from the surveyed ecological studies. Bars represent the average percentages for each bin realized after 1000 perturbed resamples, while grey bars indicate the 95\% confidence interval. }
\label{afoto1}
\end{figure}

In the temporal dimensions, 36\% of observations were not repeated (Fig. 1C), 15\% were repeated at short intervals (sub-second to daily), 21\% at daily to monthly intervals, 19\% at monthly to yearly intervals, and 8\% at yearly to decadal intervals. The temporal duration of studies (the total amount of time the ecological feature was directly observed, calculated as the number of repeat observations multiplied by the effective sampling duration; SI) was less than one day for 71\% of sampled observations, between one day and one month for 17\% of observations, between one month and one year for 7\% of observations, and greater than one year for 5\% of observations (including several paleoecological studies covering centuries to millennia; Fig. 1D).

Juxtaposing the observational dimensions provides further insight into the spatio-temporal distribution of ecological observations and the domains in which they are concentrated (Fig. 2). Contrasting spatial resolution with sampling interval reveals that the majority of temporally replicated observations (the 36\% that were unreplicated were excluded because they have no interval value) had resolutions of 10 cm$^2$-1 m$^2$ and were revisited at daily to yearly intervals (Fig. 2A). A less dense, oblong concentration of observations bounded on the lower right by monthly to yearly observations at 100 m$^2$ and on the upper left by near-daily to monthly observations with 1-10 ha resolution is also evident. This lower right to upper left orientation reflects the tradeoff between resolution and return interval that is typical of satellite imaging \cite{estes_platform_2016}, and stands in contrast to the upper right to lower left line that stretches between this concentration and the high frequency (minute-hour intervals), high spatial resolution (0.1-100 cm$^2$) observations. This line demonstrates the opposite tradeoff that occurs with field-based observations, where larger plot sizes demand greater effort that in turn reduces sampling frequency \cite{kareiva_spatial_1988}.

@@

\begin{figure}[!ht]
%\begin{wrapfigure}{c}{1\textwidth}
\includegraphics[width=1\textwidth]{figures/kde45.pdf}
\vspace{-0.15 cm}
\caption{Two-dimensional kernel density estimates of observational densities within the domains defined by A) sampling interval and spatial resolution, B) temporal duration and spatial extent, C) spatial resolution and spatial extent, and D) sampling interval and temporal duration. Density estimates were applied to the log-transformed values of each observational dimension, and density estimates are rescaled to represent percentages. }
\label{afoto1}
\end{figure}

Contrasting temporal duration and spatial extent (for all observations) reveals that the bulk of ecological observations covered total time periods of one minute to $<$1 month and areas of $\leq$1 m$^2$ (Fig. 2B). The next greatest concentration were instantaneous ($\leq$1 second) observations of $<$1 m$^2$ extent, followed by a fainter concentration centered on 1 hour duration and between 1 m$^2$ and 1 ha in extent. Very few observations had spatial extents $>$1,000,000 ha, and because these were collected by satellites that make near-instantaneous measurements, they tended to have aggregate durations of $\leq$1 hour. In contrast, the longest periods of observation were provided by paleo-ecological or long-term weather stations, both of which produce continuous, point-scale measurements (e.g. sediment cores and weather instrumentation).

Comparing the two spatial dimensions against one another (for all observations) reveals a log-linear relationship wherein resolution strongly constrains extent, such that the ratio between extent and resolution falls as the latter increases (Fig. 2C). As with interval, this pattern indicates that spatial replication diminishes with plot size \cite{kareiva_spatial_1988} in field-based studies, which accounted for 80\% of observations. This stands in contrast to the 6.5\% of observations that were made with remote sensing, which have fine spatial resolutions relative to their large extents. These comprise the majority of observations in the four low density patches (e.g. 0.1-10 ha resolution and 1-10 million ha extent) falling above the denser line that terminates at 100 ha resolution and 1000 ha extent.

A similar linear relationship between sampling interval and temporal duration was not evident (unreplicated observations were excluded because they lacked interval values). For repeated observations, the greatest concentration was among those made at daily to yearly intervals with overall durations of hours to months (Fig. 2D). A second concentration of sub-monthly to yearly observations lasting one second or less is also evident. Very few repeat observations exceeded a year in duration, and these were collected at daily to yearly intervals.

>>>>

We calculated these dimensions from 379 discrete observations reported within a 134 paper subset of 348 randomly selected articles (from 42,918 total). An additional 62 papers that were cited as the source of observations in the selected papers were also reviewed. We confined our analysis to observations made of ``natural'' \cite{tilman_ecological_1989}, or non-experimentally manipulated, systems, given that the inclusion of experiments could have skewed our assessments towards the relatively fine scales at which such studies typically focus.

@@

In terms of resolution, the majority (67\%) of observations were collected in plots of $<$1 m$^2$ resolution, 24\% were collected within plots of 1 m$^2$ up to 1 ha, and the remaining 9\% in plots of $\geq$1 ha (Fig. 1A). The extent of 19\% of observations was $<$10 ha, 23\% covered 10-1,000 ha, 42\% 1000-1,000,000 ha, and 16\% $>$1,000,000 ha (Fig. 1B).

@@

\begin{figure}[!ht]
%\begin{wrapfigure}{c}{1\textwidth}
\includegraphics[width=1\textwidth]{figures/hists_bs.pdf}
\vspace{-0.15 cm}
\caption{Histograms of the resolution (A), extent (B), interval (C), and duration (D) of observations collected from the surveyed ecological studies. Bars represent the average percentages for each bin realized after 1000 perturbed resamples, while grey bars indicate the 95\% confidence interval. }
\label{afoto1}
\end{figure}

In the temporal dimensions, 37\% of observations were not repeated (Fig. 1C), 17\% were repeated at short intervals (sub-second to daily), 20\% at daily to monthly intervals, 18\% at monthly to yearly intervals, and 8\% at yearly to decadal intervals. Duration was one day or less for 31\% of sampled observations, while 10\% covered one day to one month, 23\% lasted one month to one year, 27\% covered 1-10 years, and 10\% spanned a decade or more (including several paleoecological studies covering centuries to millennia; Fig. 1D).

Juxtaposing these observational dimensions provides further insight into the spatio-temporal distribution of ecological observations and the domains in which they are concentrated (Fig. 2). Contrasting resolution with interval reveals that the majority of temporally replicated observations (the 37\% that were unreplicated were excluded because they lack interval values) had resolutions of 10 cm$^2$-1 m$^2$ and were revisited at daily to yearly intervals (Fig. 2A). A less dense, oblong concentration of observations bounded on the lower right by monthly to yearly observations at 100 m$^2$ and on the upper left by near-daily to monthly observations with 1-10 ha resolution is also evident. This lower right to upper left orientation reflects the tradeoff between resolution and interval that is typical of satellite imaging \cite{estes_platform_2016}, and stands in contrast to the upper right to lower left line that stretches between this concentration and the high frequency (minute-hour intervals), high spatial resolution (0.1-100 cm$^2$) observations. This line demonstrates the opposite tradeoff that occurs with field-based observations, where larger plot sizes demand greater effort that in turn reduces sampling frequency \cite{kareiva_spatial_1988}.

@@

\begin{figure}[!ht]
%\begin{wrapfigure}{c}{1\textwidth}
\includegraphics[width=1\textwidth]{figures/kde45.pdf}
\vspace{-0.15 cm}
\caption{Two-dimensional kernel density estimates of observational densities within the domains defined by A) interval and resolution (of temporally replicated observations only), B) duration and extent, C) resolution and extent, and D) interval and duration (of temporally replicated observations). Density estimates were applied to the log-transformed values of each observational dimension, and density estimates are rescaled to represent percentages.}
\label{afoto1}
\end{figure}

Contrasting duration and extent (for all observations) reveals two primary domains of observational concentration. The first consists of observations spanning one month to one decade in time and 10-1000 ha in space, while the second is defined by observations of one year to several decades that cover 10,000 to 1,000,000 ha (Fig. 2B).  Three other notable, but lesser areas of concentration are also evident, including small area observations (0.1-1 ha) covering one month to decade, and short duration, temporally unreplicated observations ($<$1 day) of either 1-10 ha or 10,000-1,000,000 ha.

% the bulk of ecological observations span one month to one decade in time and covered total time periods of one minute to $<$1 month and areas of $\leq$1 m$^2$ (Fig. 2B). The next greatest concentration were instantaneous ($\leq$1 second) observations of $<$1 m$^2$ extent, followed by a fainter concentration centered on 1 hour duration and between 1 m$^2$ and 1 ha in extent. Very few observations had spatial extents $>$1,000,000 ha, and because these were collected by satellites that make near-instantaneous measurements, they tended to have aggregate durations of $\leq$1 hour. In contrast, the longest periods of observation were provided by paleo-ecological or long-term weather stations, both of which produce continuous, point-scale measurements (e.g. sediment cores and weather instrumentation).

Comparing the two spatial dimensions against one another (for all observations) shows a primary concentration of observations with 10 cm$^2$ to 100 m$^2$ resolution that have extents ranging between slightly over 1,000 to nearly 1,000,000 ha (Fig. 2C). The second-most prominent concentration consists of higher resolution (1 cm$^2$-1 m$^2$), smaller extent (10-1000 ha) observations, beneath which lies a third and fainter concentration of 1-1,000 cm$^2$ resolution, 1000 m$^2$ to $<$10 ha.  These three concentrations suggest a tendency for observational extent to increase with resolution, which is a relationship that becomes more pronounced in the (less densely observed) portion of the domain where resolutions $\geq$100 m$^2$.

%reveals a log-linear relationship wherein resolution strongly constrains extent, such that the ratio between extent and resolution falls as the latter increases (Fig. 2C). As with interval, this pattern indicates that spatial replication diminishes with plot size \cite{kareiva_spatial_1988} in field-based studies, which accounted for 80\% of observations. This stands in contrast to the 6.5\% of observations that were made with remote sensing, which have fine spatial resolutions relative to their large extents. These comprise the majority of observations in the four low density patches (e.g. 0.1-10 ha resolution and 1-10 million ha extent) falling above the denser line that terminates at 100 ha resolution and 1000 ha extent.

A similar tendency for duration to increase with interval was also evident amongst temporally replicated observations (Fig. 2D), where the majority of observations were repeated at daily to decadal intervals and spanned  $\geq$ 1 month to $<$100 years.  The orientation of this concentration shows that observations lasting one year to one decade tend to have corresponding intervals, while those lasting one month to one year have daily to monthly intervals. This suggests that most long-duration observations have just a single temporal replicate. The fainter densities of observations having sub-daily intervals shows that relatively few high frequency, long duration ecological measurements are undertaken.

%A similar linear relationship between sampling interval and temporal duration was not evident (unreplicated observations were excluded because they lacked interval values). For repeated observations, the greatest concentration was among those made at daily to yearly intervals with overall durations of hours to months (Fig. 2D). A second concentration of sub-monthly to yearly observations lasting one second or less is also evident. Very few repeat observations exceeded a year in duration, and these were collected at daily to yearly intervals.
